### 2020-09-11 Analysis/Code/Paper.Rmd
tracing the bug

# Empirical Model of Demand

@@

# Discounting and Review Bin Transitions
\label{descriptiveEvidence}

## Stylized Model

Consider a firm that has a review status $s \in (0, 1)$. A firm can give a discount at cost $c$ that can probabilistically change its status to exceed 1 (an upgrade), to go below 0 (a downgrade), or to remain within the $(0,1)$ band (no change). The result depends on whether the additional customers who buy the product on a discount happen to leave more positive or more negative reviews, and on the  weight of hte new reviews relative to the existing ones. Assume that with probability $p$ the firm's status becomes $s+x$, and with probability $1-p$ it becomes $s-x$. Firm's utility $u(s)$ from having a review status $s \in (0,1)$ is 0, while it earns a utility $u^H$ from the upgraded status and a utility $u^L < 0 < u^H$ from the downgraded status. The expected utility from giving a discount is 
\begin{equation}
  \EE{U(s+X)} = p U(s + x) + (1-p) U(s-x) - c
\end{equation}

When would the firm be willing to give a discount? Consider a firm with $s > 1/2$, i.e., the one that is close to a review upgrade. For that firm the expected value of giving a discount is 
$$
  \EE{U(s+X)} = 
  \begin{cases}
  -c & x < 1-s\\
  p u^H -c & x \in [1-s, s) \\
  pu^H + (1-p)u^L - c & x \ge s
  \end{cases}
$$
\noindent The firm would choose to give a discount if $\EE{U(s+X)}  > 0$. A trivial comparative statics exercise tells us that the firm is more likely to do so if the probability $p$ to get a positive shock to the review status if high, if the payoff $u^H$ of upgrading the status is high, and if the loss from downgrading the status $-u^L$ is low.

What is interesting about this problem, is that the probability of improving the review status $p$ does not have to exceed $1/2$ for the firm to give a discount, i.e., the reviews it gets on a discount could be worse than the reviews it receives off a discount. Indeed, if the potential change to the score is moderate ($x \in [1-s, s]$), then a successful outcome leads to a transition, while an unsuccessful outcome leaves the firm with the current review status. In this case the firm is not risking anything, and is willing to “pull the plug” simply to gamble on the positive outcome. Even if the potential change to the review status is substantial $(x \ge s)$, the firm could still find  it profitable to give a discount and risk the deterioration of its review status if the benefits of having an improved status are large enough. Similar conclusions follow for the case of $s < 1/2$, when a firm is closer to a negative transition.

The main takeaway from this stylized model is that the relationship between proximity to transitions and discounting behavior is theoretically ambiguous. In the data we can expect to find both firms that are more likely to give a discount when they are close to a transition (gambling to upgrade their status, or trying to escape from the downgrading threshold), and firms that are less likely to discount when close to a transition (the downside risk is too big). I will now investigate this relationship in the data, and will use the intuition developed here to interpret the results.

## Empirical Analysis

@@

The departing point of the analysis is Figure \ref{discountsTransitions}, which shows that transitions between review labels are often preceded by discounts. The graph shows that two weeks prior to a transition only `r transPanel[x==-14, round(mean(100*(discount > 0)))]`% of games were on a discount, essentially the sample average, but that this number more than doubles to around `r transPanel[x==-1, round(mean(100*(discount > 0)))]`% one day before the label change. Regression analysis controlling for review label, number of reviews, age, day of the week and week time effects, as well as time since the previous discount, confirms that the association depicted in Figure \ref{discountsTransitions} is robust (see Table \ref{transitionReg} in the Appendix). A game is approximately `r round(reg1$coefficients[2]*14*100)` pp more likely to be on a discount on the day of the transition than it is two weeks prior to it, and it is `r round(-reg1$coefficients[3]*14*100)` pp less likely to be on a discount two weeks following the transition. Given that the probability for a game to be on a discount on a random day is `r round(panel[, mean((discount > 0), na.rm=T)]*100)`% in the sample, these effects amount to `r round(100*(reg1$coefficients[2]/panel[, mean((discount > 0), na.rm=T)]))`% and `r round(100*(-reg1$coefficients[3]/panel[, mean((discount > 0), na.rm=T)]))`% changes in the daily probability of a discount, which is quite sizable.

@@

# define how many positive and negative reviews are needed to transition, manually
# for every threshold
R.low <- 0.4
R.high <- 0.7
panel[score < R.high*100 & score >= R.low*100,
      needPos := ceiling((R.high*reviews-pReviews)/(1-R.high))]
panel[score < R.high*100 & score >= R.low*100,
      needNeg := floor(abs((R.low*reviews-pReviews)/(1-R.low)))+1]
# corrects for the rounding error caused by machine precision division of, say, 4/10 != 0.4. If you are supposed to be brought exactly to the cutoff, your needNeg is one review more than that
panel[(pReviews - needNeg)/(reviews-needNeg) == R.low & needNeg > 0,  needNeg := needNeg + 1]

@@

star.out <- stargazer(reg1, reg2, reg3,
             label = "potentialTransitionsResults",
             table.placement = "hp",
             header = FALSE,
             title = "Discounts Close to Potential Transitions",
             se=list(se1, se2, se3),
             dep.var.labels = c("Discount Probability"),
             column.labels = c("Full", "1 W. to Tr.", "Simple Tr."),
             covariate.labels = c("Close to Pos. Transition", "Close to Neg. Transition",
                                  "Log Days to Pos. Tr.", "Log Days to Neg Tr.",
                                  "Log Reviews", "Score", "Mostly Positive",
                                  "Positive", "Very Positive",
                                  "Age", "Age $\\le 14$"),
             model.numbers = F,
             omit = c("day", "week", "tWODisc"),
             omit.stat=c("f","adj.rsq","ser"),
             no.space=T)

@@

>>>>

@@

# Descriptive Evidence
\label{descriptiveEvidence}

In this section I present evidence that firms use discounting to manage their online reputation, in particular, by running price promotions when they are close to improving their review label. For now, let us resort to a very simplistic story of why a firm would be interested in doing that. Assume that discounts increase the inflow customers, and that more customers means more reviews. Also, take for granted, for now, that better review labels improve sales. Then, if a firm is one positive review away from upgrading its review label, it might be tempted to speed things up by giving a discount. Similarly, a product that is close to sliding into a worse review bin, could be tempted to run a price promotion, in order to avoid a slump in its reviews. I will address this story in more detail in the discussion part of this section, but for now I turn attention to simply establishing that this sort of behavior indeed takes place in the sample.

@@

The departing point of the analysis is Figure \ref{discountsTransitions}, which shows that transitions between review labels are very often preceded by discounts. The graph shows that two weeks prior to a transition only 15% of games were on a discount, a number close to the sample average, but that this number more than doubles to around 35% one day before the label change. Regression analysis controlling for review label, number of reviews, age, day of the week and week time effects, as well as time since the previous discount, confirms that the association depicted in Figure \ref{discountsTransitions} is robust (see Table \ref{transitionReg} in the Appendix). A game is approximately `r round(reg1$coefficients[2]*14*100)` pp more likely to be on a discount on the day of the transition than it is two weeks prior to it, and it is `r round(-reg1$coefficients[3]*14*100)` pp less likely to be on a discount two weeks following the transition. Given that the probability for a game to be on a discount on a random day is `r round(panel[, mean((discount > 0), na.rm=T)]*100)`% in the sample, these effects amount to `r round(100*(reg1$coefficients[2]/panel[, mean((discount > 0), na.rm=T)]))`% and `r round(100*(-reg1$coefficients[3]/panel[, mean((discount > 0), na.rm=T)]))`% changes in the daily probability of a discount, which is quite sizable.

@@

# define how many positive and negative reviews are needed to transition, manualy
# for every threshold
R.low <- 0.4
R.high <- 0.7
panel[score < R.high*100 & score >= R.low*100,
      needPos := ceiling((R.high*reviews-pReviews)/(1-R.high))]
panel[score < R.high*100 & score >= R.low*100,
      needNeg := floor(abs((R.low*reviews-pReviews)/(1-R.low)))+1]
# corrects for the rounding error caused by machine precision division of, say, 4/10 != 0.4. If you are supposed to be brought exactly to the cutoff, your needNeg is one review more than that
panel[(pReviews - needNeg)/(reviews-needNeg) == R.low & needNeg > 0,  needNeg := needNeg + 1]

@@

star.out <- stargazer(reg1, reg2, reg3,
             label = "potentialTransitionsResults",
             table.placement = "hp",
             header = FALSE,
             title = "Discounts Close to Potential Transitions",
             se=list(se1, se2, se3),
             dep.var.labels = c("Discount Probability"),
             column.labels = c("Full", "1 W. to Tr.", "Simple Tr."),
             covariate.labels = c("Close to Pos. Transition", "Close to Neg. Transition",
                                  "Log Days to Pos. Tr.", "Log Days to Neg Tr.",
                                  "Mostly Positive", "Positive", "Very Positive",
                                  "Score", "Log Reviews", "Age", "Age $\\le 14$"),
             model.numbers = F,
             omit = c("day", "week", "tWODisc"),
             omit.stat=c("f","adj.rsq","ser"),
             no.space=T)

@@

# Empirical Model
